As the economy develops and travel demand grows, motorcycles, valued for their convenience, are becoming more numerous. Concurrently, due to tougher traffic management and some riders' lax safety awareness, motorcycle - related accidents are on the rise. Helmet use significantly impacts rider safety in such incidents. This paper presents a real - time helmet - wearing detection system for motorcycle riders. It streamlines detection, visualizes results, minimizes manual input, and boosts law enforcement efficiency, covering model training and system development.

In terms of detection model training, based on the Yolo target recognition algorithm, this paper constructs a dual model collaborative architecture: first, it focuses on rapid recognition of the overall helmet wearing status of motorcycles and their riders in images or videos; Second, the key area of the former output is the input, which can accurately locate and identify the driver ID. in the process of research, the training effects of three different scale models, yoov11n, yoov11s and yoov11m, are systematically compared and analyzed. Through continuous optimization of parameters, the accuracy and stability of helmet detection and driver ID detection are improved.

In terms of system development, this paper develops a front-end and back-end separated system based on Browser/Server (B/S) architecture, uses Vue framework to build B-end pages, and springboot framework to build S-end servers. The front-end browser provides users with two functions: helmet wearing detection and history query. The helmet wearing detection function allows users to customize the detection model, IOU and confidence, and supports browsing and saving the detection results. The history query function supports users' filtering of query fields and data visualization. The back-end server is also divided into two modules, target detection and record query, which are respectively used to connect the two functions of the front-end. The target detection module receives the data input by the user, calls the model for detection, and finally returns the detection results. The record query module queries the history records in the database according to the filter conditions entered by the user and returns them.